Suplementary Code 4: Draw Interbirth Interval Predictons on the Last Birth Omited Data

### 2024-01-09 S5_response_to_reviewers.Rmd

```{r}
#compared to twinR original add data on :
#-age at first birth (AFB)
#-quantiles range mother birth year
#-quantiles range offspring birth year
#-mean + SE total births
all_tbls <- rbind(build_data_summary.table(data_births_monthly),
                  build_data_summary.table(data_births_monthly_not_last),
                  build_data_summary.table(data_births_monthly_EE),
                  build_data_summary.table(data_births_monthly_EE_not_last))
```

## Comment 1: Logistic regression

We can test the hypothesis that mothers are statistically more likely to stop having children after a twin birth by fitting a logistic regression model with the response variable being the binary variable *last* and the predictor variable being the binary variable *twin*.

```{r}
#fit logistic regression Estonia
logit_fit_EE <- glm(last ~  twin,
                 data = data_births_monthly_EE,
                 family = binomial(link = "logit"))
summary_logit_fit_EE <-  summary(logit_fit_EE)
summary_logit_fit_EE
```

```{r}
#fit logistic regression Other European
logit_fit <- glm(last  ~  twin,
                 data = data_births_monthly,
                 family = binomial(link = "logit"))
summary_logit_fit_W <-  summary(logit_fit)
summary_logit_fit_W
```

## Comment 2: Proportion of twins

@@

```

```{r}
#compare the precentage of twin births between last births and not last births
cat("Western dataset:\n")
tblw <- table(data_births_monthly[c("twin", "last")])
tblw
cat("\n")
#now calcualte the % by hand
cat("Last twins: ", round(100 * tblw[2,2] / sum(tblw[,2]), 2), "%\n")
cat("Prior twins: ",round(100 * tblw[2,1] / sum(tblw[,1]), 2), "%\n")

```

#### Test the Population Interaction

We check if the population has an effect in a reversed model where the predictor variable is the binary variable *last* and the response variable is the binary variable *twin*.

@@

    geom_ribbon(data=fig2_EE_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    color=estLab, fill = estLab),
                alpha = 0.3) + 
    geom_line(data=fig2_orig_plot_data,
             aes(y = estimate,  x=births_total, color=westLab),  size = 1) +
  geom_point(data=dmm_orig[dmm_orig$births_total <19, ],
               aes(x=births_total, y=prob_twin, color=estLab, fill = estLab),
               alpha=0.1,
               position = position_jitter()) +
    stat_summary(data=dmm_orig[dmm_orig$births_total <19, ],
                 aes(x=births_total, y=prob_twin,
                     color=westLab, fill=westLab),
                 alpha=1,
                 shape = 1,
                 position = position_nudge(x = 0.1),
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_orig_plot_data,
                aes(y = estimate,  x=births_total, ymin = lwr, ymax = upr,
                    fill=westLab),
                alpha = 0.1) +
    ggplot2::scale_x_continuous(breaks = 1:18) +

    ggplot2::coord_cartesian() +
    labs(subtitle = paste0("(a) All births, ", infoTxt),
         y="Per-birth twin. prob.",
         x="Maternal total births")
p2 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") + labs(fill = "population")

p2
```

The plot illustrates that we cannot plot the raw data along with model predictions, as such figure would be un-informative and highly confusing. We cannot think of a way about how to put such an amount of semi-categorical data into a plot so that it would give additional information.

## Comment 4: Opposite-sex Twins

We can do the same analysis for opposite-sex  twins.

@@

```{r}
dmm_EE_mixed_nl_plot <- dmm_EE_mixed_nl
dmm_EE_mixed_nl_plot$births_total <- dmm_EE_mixed_nl_plot$births_total + 1
```

```{r}
#colors
bc <- c("purple", "black")
fig2_ext_orig <- ggplot() +  
  geom_line(data=fig2_EE_plot_data,
              aes(y = estimate,  x=births_total, color="EE all"),
            linewidth = 1) +
    stat_summary(data=dmm_EE_mixed[dmm_EE_mixed$births_total <14, ],
                 aes(x=births_total, y=prob_twin, color="EE all",
                     fill = "EE all"), alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_EE_plot_data,
                aes(y = estimate,  x=births_total,
                    ymin = lwr, ymax = upr, color="EE all", fill = "EE all"),
                alpha = 0.3) + 
    geom_line(data=fig2_EE_plot_data_nl,
              aes(y = estimate,  x=births_total, color="EE w/o last"),
              linewidth = 1) +
    stat_summary(data=dmm_EE_mixed_nl_plot[dmm_EE_mixed_nl_plot$births_total <14,],
                 aes(x=births_total, y=prob_twin, color="EE w/o last",
                     fill = "EE w/o last"), alpha=0.5,
                 fun.data=mean_se) +
    geom_ribbon(data=fig2_EE_plot_data_nl,
                aes(y = estimate,  x=births_total, ymin = lwr,
                    ymax = upr, color="EE w/o last", fill = "EE w/o last"),
                alpha = 0.3) + 
    ggplot2::scale_x_continuous(breaks = 1:18) + 
    ggplot2::scale_y_continuous(breaks = seq(0,0.03, by=0.002)) +
    ggplot2::coord_cartesian() +
    labs(subtitle = "Prediction only mixed sex twins + data mean with SE",
         y="Per-birth twin. prob.",
         x="Maternal total births")
p2 <- fig2_ext_orig + base_theme(larger=8) + scale_color_manual(values=bc) + 
  scale_fill_manual(values=bc) + guides(color="none") + labs(fill = "dataset")

p2 
```

The mixed sex twinning rate is independent on the number of total births if last births are excluded from the data. If last births are included, the association between mixed sex twinning rate and number of total births is negative.
